---
title: "Data Analysis Homework 1"
author: "Artin Tajdini"
date: "February 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

### n = 4 table

```{r}
n = 4

#algoritm for n = 4k
m4 = matrix(1:n^2, nrow = n, ncol = n, byrow = TRUE)

m4[1:n/4, 1:n/4] = n^2 + 1 - m4[1:n/4, 1:n/4]
m4[1:n/4, (3*n/4 + 1):n] = n^2 + 1 - m4[1:n/4, (3*n/4 + 1):n]
m4[(3*n/4 + 1):n, 1:n/4] = n^2 + 1 - m4[(3*n/4 + 1):n, 1:n/4]
m4[(3*n/4 + 1):n, (3*n/4 + 1):n] = n^2 + 1 - m4[(3*n/4 + 1):n, (3*n/4 + 1):n]
m4[(n/4 + 1):3*n/4, (n/4 + 1):3*n/4] = n^2 + 1 - m4[(n/4 + 1):3*n/4, (n/4 + 1):3*n/4]

m4

```

the algorithm above generates magic matrix for every n = 4k

### n = 5 table

```{r}
n = 5

#algoritm for n = 2k + 1
magic.odd = function(n){
  m5 = matrix(0, nrow = n, ncol = n)
  cnt = 1
  i = 1 + (n/2)
  j = n
  
  while (cnt <= n * n) {
    if (i == 0 & j == n + 1) {
      j = n - 1
      i = 1
    } else{
      j = 1 + ((j - 1 + n) %% n)
      i = 1 + ((i - 1 + n) %% n)
    }
    
    if( m5[i,j] != 0 ){
      j = j - 2
      i = i + 1
    } else{
      m5[i,j] = cnt
      cnt = cnt + 1
      i = i - 1
      j = j + 1
    }
    
  }
  m5 
}


m5 = magic.odd(5)
m5

```

the algorithm above generates a magic matrix for every odd n

### n = 6 table

```{r}
n = 6

#algoritm for n = 4k + 2
mbase = magic.odd(n/2)

#flipping magic base
for (i in 1:(n/4)) {
  tmp = mbase[,i]
  mbase[,i] = mbase[,((n/2)-i) + 1]
  mbase[,((n/2) - i) + 1] = tmp
}

#rotating magic base

for(i in 1:((n/4))){
  for (j in 1:((n/4) + 1)) {
      tmp = mbase[j, (n/2) - i + 1]
      mbase[j, (n/2) - i + 1] = mbase[i,j]
      mbase[i,j] = mbase[(n/2) - j + 1, i]
      mbase[(n/2) - j + 1, i] = mbase[(n/2) - i + 1, (n/2) - j + 1]
      mbase[(n/2) - i + 1, (n/2) - j + 1] = tmp

  }
}

#main part
m6 = matrix(0, nrow = n, ncol = n, byrow = TRUE)

m6[1:(n/2), 1:(n/2)] = mbase
m6[(n/2 + 1):n, 1:(n/2)] = mbase + 3 * n^2/4
m6[1:(n/2), (n/2 + 1):n] = mbase + 2 * n^2/4
m6[(n/2 + 1):n, (n/2 + 1):n] = mbase + n^2/4


lc = floor((n/4)) + 1
rc = lc - 1

for (i in 1:(n/2)) {
  for (j in 1:(n/2)) {
    if(j < lc | (i == lc & j == lc)){
      if(!(j == 1 && i == lc )){
        tmp = m6[i, j]
        m6[i,j] = m6[i + (n/2), j]
        m6[i + (n/2), j] = tmp
      }
      
    }
  }  
}

m6

```

## Question 2

first I find every valid state for first 5 numbers and put it in a list of lists named p1!

then I find every valid state for second 5 numbers based on every valid state of first 5 and put this valis states in a list of list named p2!

at last I fond every valid state of last three numbers base on every valid state of first 10 

```{r}
a = numeric(13)

#first part

p1 = vector(mode = "list")

for (i in 1:13) {
  if(i %in% a)
    next
  a[1] = i
    for (j in 1:13) {
      if(j %in% a)
        next
      a[2] = j
      
        for (k in 1:13) {
          if(k %in% a)
            next
          a[3] = k
          
          for (l in 1:13) {
            if(l %in% a)
              next
            a[4] = l
            
            for (z in 1:13) {
              if(z %in% a)
                next

              if( i / (j + k) == (j + k) / (k + l + z)
                  & (i + j) / (j + k) < 1 ){
                    p1 = append(p1, list( list(i,j,k,l,z) ) )
              }
            }
          a[4] = 0
        }
        a[3] = 0
      }
  
    a[2] = 0
  }
  
  a[1] = 0
}

#second part

p2 = vector(mode = "list")

for (e in p1) {
  a[1:5]= as.numeric( e )
  
  for (i in 1:13) {
  if(i %in% a)
    next
  a[6] = i
  for (j in 1:13) {
    if(j %in% a)
      next
    a[7] = j
    
    for (k in 1:13) {
      if(k %in% a)
        next
      a[8] = k
      
      for (l in 1:13) {
        if(l %in% a)
          next
        a[9] = l
        
        for (z in 1:13) {
          if(z %in% a)
            next
          
          if( i / (j + i) == (j + i) / (k + l + j)
              &  i / (j + i) == (k + l + j) / ( a[5] + l + z)){
            p2 = append(p2, list( c(e, list(i,j,k,l,z)) ))
          }
        }
        a[9] = 0
      }
      a[8] = 0
    }
    
    a[7] = 0
  }
  
  a[6] = 0
}
}

#last part

ans = numeric(13)

for (e in p2) {
  a[1:10]= as.numeric( e )
  
  for (i in 1:13) {
    if(i %in% a)
      next
    a[11] = i
    for (j in 1:13) {
      if(j %in% a)
        next
      a[12] = j
      
      for (k in 1:13) {
        if(k %in% a)
          next
        a[13] = k
        
            if( (i + j) / (j + k) == (j + k) / (k + a[10])
                & (i + j) / (j + k) < 1){
              ans = a
        }
        a[13] = 0
      }
      
      a[12] = 0
    }
    
    a[11] = 0
  }
}

ans

```

## Third Question

for every part I check every $2^n$ possibilities of people on circle being liar and if state is valid I add number of liars on that state to a list

### part 1

```{r}
number.lyer1 = numeric()

for (i in 0:((2^16) - 1)) {
  arr = as.integer( intToBits(i) )
  flag = FALSE
  for (j in 1:16) {
    l = j - 1
    r = j + 1
    
    if(l == 0)
      l = 16
    
    if(r == 17)
      r = 1
    
    if(arr[j] == 0 & sum(arr[c(l,r)]) != 2)
      flag = TRUE
    
    if(arr[j] == 1 & sum( arr[c(l,r)] ) == 2)
      flag = TRUE
    
  }
  
  if(!flag)
    number.lyer1 = append(number.lyer1, sum(arr))
  
}

unique(number.lyer1)

```

### part 2

```{r}
number.lyer2 = numeric()

for (i in 0:((2^12) - 1)) {
  arr = as.integer( intToBits(i) )
  flag = FALSE
    for (j in 1:12) {
      l = j - 1
      r = j + 1
      
      if(l == 0)
        l = 12
      
      if(r == 13)
        r = 1
      
      if(arr[j] == 0 & sum(arr[c(l,r)]) != 1)
        flag = TRUE

      if(arr[j] == 1 & sum( arr[c(l,r)] ) == 1)
        flag = TRUE

    }
    
    if(!flag)
      number.lyer2 = append(number.lyer2, sum(arr))
}

unique(number.lyer2)

```

### part 3

```{r}
number.lyer3 = numeric()

for (j in 0:((2^8) - 1)) {
  brr = as.integer( intToBits(j) )

  if( sum(brr) == 4 ){
    for(i in 0:((2^8) - 1)){
      arr = as.integer( intToBits(i) )
      
      
      flag = FALSE
      for (k in 1:8) {
        l = k - 1
        r = k + 1
        
        if(l == 0)
          l = 8
        
        if(r == 9)
          r = 1
        
        if(arr[k] == 0 & brr[k] == 0 & sum(arr[c(l,r)]) != 1)
          flag = TRUE
        
        if(arr[k] == 1 & brr[k] == 0 & sum( arr[c(l,r)] ) == 1)
          flag = TRUE
  
        if(arr[k] == 0 & brr[k] == 1  & sum(arr[c(l,r)]) != 2)
          flag = TRUE
        
        if(arr[k] == 1 & brr[k] == 1 & sum( arr[c(l,r)] ) == 2)
          flag = TRUE
        
      }
      
      if(!flag)
        number.lyer3 = append(number.lyer3, sum(arr))
    }
  }
}

unique(number.lyer3)

```